---
title: "GoodReads"
author: "Team 5"
date: "28/05/2020"
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
---


<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
#importing libraries and dataset
library(readxl)
library(sqldf)
library(dplyr)
library(ggplot2)
library(corrplot)
library(tidyr)
library(corrgram)
library(RColorBrewer)
library(dplyr)
setwd("C:/Users/Tina/Desktop/STT/Project/Data")
books=read_excel('Cleaned_data_final_v1.xlsx')
options(scipen = TRUE)
books$X=NULL
books$X.1=NULL
books$...1=NULL

```

```{r}
names(books)
```

## Subsetting 

In this section, we remove observations based on the following conditions

1. Books with number of pages greater than 1200

2. Books with number of ratings count less than 50

3. Books with author as 'Anonymours'

```{r}
books<-sqldf('select * from books where num_pages<1200 and ratings_count>50') 
books<-sqldf('select * from books where Awards<20')
books<-sqldf("select * from books where primary_author!='Anonymous'")
```

## Binning

In this section we bin the cateforical variables based on the following conditions

1. Publishers with less than 20 observations are binned as 'Others'

2. Genre with less than 20 observations are binned as 'Others'

3. Languages with less than 5 observation are binned as 'Others'

```{r}
dist<-sqldf('select count(bookID) as count, publisher from books  group by publisher having count>=20')
genre<-sqldf('select count(bookID) as count, Primary_Genre from books  group by Primary_Genre having count>=20')
lang<-sqldf('select count(bookID) as count, language_code from books group by language_code having count > 5')

for (i in 1:length(books$publisher) ){
  if(!(books$publisher[i] %in% dist$publisher)){
    books$publisher[i]='Others'
  }
}


for (i in 1:length(books$Primary_Genre) ){
  if(!(books$Primary_Genre[i] %in% genre$Primary_Genre)){
    books$Primary_Genre[i]='Others'
  }
}

for (i in 1:length(books$language_code) ){
  if(!(books$language_code[i] %in% lang$language_code)){
    books$language_code[i]='Others'
  }
}

```

## Data Transformation
In this section we transform the data under follwoing conditions

1. Normalize the no of followers of the author

2. Factorize all the categorical variables 

```{r}
normalize <- function(x){
  return((((x-min(x))/ max(x)-min(x))) * 500)
}
books$normal_auth<-normalize(books$Auth_followers)
#Converting some columns to categorical
books$Series<-as.factor(books$Series)
books$publisher<-as.factor(books$publisher)
books$Primary_Genre<-as.factor(books$Primary_Genre)
books$language_code<-as.factor(books$language_code)
```

## Exploratory Data Analysis
```{r}
# distribution
ggplot(data=books)+
  geom_freqpoly(aes(x=average_rating),color="black")+labs(title='Avergae Rating Distribution')

#distribution vs Series
ggplot(data=books)+
  geom_freqpoly(aes(x=average_rating,color=Series),position="identity")

# NO of pages
ggplot(data=books, mapping = aes(x=average_rating,y=num_pages)) + geom_point() + geom_smooth(method="lm") + labs(title="Average Ratings Vs No Of Pages")

# No of Ratings
ggplot(data=books, mapping = aes(x=ratings_count,y=num_pages)) + geom_point() + geom_smooth(method="lm") + labs(title="Average Ratings Vs Rating Count")

#Auth followers
ggplot(data=books, mapping = aes(x=ratings_count,y=Auth_followers)) + geom_point() + geom_smooth(method="lm") + labs(title="Avergae Rating Vs Author Followers")

```

## Correlation Analysis

```{r}
books_numeric <- select (books,average_rating,ratings_count,text_reviews_count, Auth_followers,Auth_books)
corrplot(corrgram(books_numeric), method="circle", type="lower", tl.col="black",tl.srt=45,col=brewer.pal(n=5, name="Set2"))
```

## Modelling
### Full Model
The following columns were considered for the full model
1. Number of Pages
2. Primary Genre
3. Awards
4. Series
5. No of Author's followers 
5. No of books for the author
6. Language code
7. Publisher

```{r}
books_model <- select(books,average_rating, num_pages, Primary_Genre, Awards, Series, normal_auth,Auth_books,language_code,publisher)
books$Series<-as.factor(books$Series)
books$Primary_Genre<-as.factor(books$Primary_Genre)
books$publisher<-as.factor(books$publisher)
books$language_code<-as.factor(books$language_code)
full=lm(average_rating ~ ., data = books_model)
summary(full)
```

### Residual Plot

```{r}
residuals <- books_model$average_rating - fitted(full)
resids <- data.frame("Fitted" = fitted(full), "Residuals" = residuals)
ggplot(data = resids, aes(x = Fitted, y = Residuals) )+
  geom_point(aes(x = Fitted, y = Residuals, alpha=0.05, color='red'),size=0.8) +
  geom_hline(aes(yintercept = 0))+ labs(title = 'Residual Plot')

ggplot(data = resids, aes(sample = Residuals)) +
  stat_qq(size=0.8) + stat_qq_line() + labs(title = 'Q-Q Plot')

```

## Backward Stepwise Feature selection with AIC
```{r}
# BACKWARD FROM FULL
options(scipen=999)
#AIC
step1 <- step(lm(average_rating ~ ., data = books_model), direction = "backward", trace = 0)
summary(step1)
```
### Residual Plot and Q-Q Plot
```{r}
residuals <- books_model$average_rating - fitted(step1)
resids <- data.frame("Fitted" = fitted(full), "Residuals" = residuals)
g1=ggplot(data = resids, aes(x = Fitted, y = Residuals) )+
  geom_point(aes(x = Fitted, y = Residuals, alpha=0.05),size=0.8) +
  geom_hline(aes(yintercept = 0)) + labs(title='RESIDUAL PLOT WITH AIC')
n1=ggplot(data = resids, aes(sample = Residuals)) +
  stat_qq() + stat_qq_line()+ labs(title='PROBABILITY PLOT WITH AIC')
```

## Backward Stepwise Feature selection with BIC
```{r}
n <- nrow(books_model)
step2 <- step(lm(average_rating ~ ., data = books_model), direction = "backward", k = log(n), trace = 0)
summary(step2)
```
### Residual Plot and Q-Q Plot
```{r}
residuals <- books_model$average_rating - fitted(step2)
resids <- data.frame("Fitted" = fitted(full), "Residuals" = residuals)

g2=ggplot(data = resids, aes(x = Fitted, y = Residuals) )+
  geom_point(aes(x = Fitted, y = Residuals, alpha=0.05,color='red'), size=0.8) +
  geom_hline(aes(yintercept = 0))  + labs(title='RESIDUAL PLOT WITH BIC')
n2=ggplot(data = resids, aes(sample = Residuals)) +
  stat_qq() + stat_qq_line() + labs(title='PROBABILITY PLOT WITH BIC')
```
## PLOTS
```{r}
library(gridExtra)
grid.arrange(g1,g2,ncol=1)
```

## Forward Stepwise Feature selection with AIC
```{r}
biggest <- formula(lm(average_rating ~ ., data = books_model))
step3 <- step(lm(average_rating ~ 1, data = books_model), direction = "forward", scope = biggest, trace = 0)
summary(step3)
```
## Forward Stepwise Feature selection with BIC
```{r}
step4 <- step(lm(average_rating ~ 1, data = books_model), direction = "forward", scope = biggest, k = log(n), trace = 0)
summary(step4)
```

## Calculating the accuracy metrics

In this section we calculate the various accuracy metrics to guide us in choosing the appropriate model.
```{r}
myANOVA <- anova(full)
MSE <- myANOVA$`Mean Sq`[9]
p<-length(step1$coefficients)
SSE_1 <- sum((step1$residuals)^2)
Cp_1 <- SSE_1/MSE-(nrow(books_model)-2*p)
PRESS_1 <- sum((resid(step1)/(1 - lm.influence(step1)$hat))^2)

p<-length(step2$coefficients)
SSE_2 <- sum((step2$residuals)^2)
Cp_2 <- SSE_2/MSE-(nrow(books_model)-2*p)
PRESS_2 <- sum((resid(step2)/(1 - lm.influence(step2)$hat))^2)

p<-length(step3$coefficients)
SSE_3 <- sum((step3$residuals)^2)
Cp_3 <- SSE_3/MSE-(nrow(books_model)-2*p)
PRESS_3 <- sum((resid(step3)/(1 - lm.influence(step3)$hat))^2)

p<-length(step4$coefficients)
SSE_4 <- sum((step4$residuals)^2)
Cp_4 <- SSE_4/MSE-(nrow(books_model)-2*p)
PRESS_4 <- sum((resid(step4)/(1 - lm.influence(step4)$hat))^2)

SSE<-c(SSE_1,SSE_2,SSE_3,SSE_4)
cp<-c(Cp_1,Cp_2,Cp_3,Cp_4)
press<-c(PRESS_1,PRESS_2,PRESS_3,PRESS_4)
aic<-c(AIC(step1),AIC(step2),AIC(step3),AIC(step4))
bic<-c(BIC(step1),BIC(step2),BIC(step3),BIC(step4))
metrics <- matrix(c(aic, bic, SSE, cp, press),ncol=5,byrow=FALSE)
colnames(metrics) <- c("AIC","BIC","SSE","CP","PRESS")
rownames(metrics) <- c("Backward AIC","Backward BIC","Forward AIC","Forward BIC")
metrics<-as.table(metrics)
metrics
```

Since, the study is more focused on interpretation of the predictor variables on the outcome variable, we choose the model with the best BIC value compromising on the R squared value.
Thus, the model chosen is the one yielded by the stepwise with BIC. (both forward and backward methods gives us the same model)

## Final Model
```{r}
summary(step2)
```


